The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-21 paper.tex
Introduced notation A(R,H,C,t) + updated fig dynamics of anthropo

\usepackage{graphicx}
\graphicspath{{figs/}}
\usepackage{tikz}
\usetikzlibrary{shapes}
%
\usepackage{amsmath} 
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
\usepackage{color, soul}
\usepackage{url}
\usepackage{multirow}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{textcomp}

@@

@@

A phenomenon that is often studied related to the human-like design of robots is
\textbf{anthropomorphism}. Besides a feeling of affinity, and empathy,
anthropomorphism can be understood as being one of the resulting (and  more or
less desired) effects of the anthropomorphic design of artificial agents (up to
the first positive peak). The essence of anthropomorphism is the human user's
perception of human-likeness in the non-human agent. Logically, anthropomorphism
is said to be enhanced with increased human-like design of the agent. There are
various other reasons and explanations for why people tend to anthropomorphize
non-human agents. In general, as anthropomorphism emerges in a (real or
imagined) interaction \citep{persson_anthropomorphism_2000} between at least two
players (the human who anthropomorphizes and the non-human agent which is
anthropomorphized) there are also two main perspectives on the phenomenon: an
artifact-centered (design) and a human-centered (cognitive, psychological) one.
The Uncanny Valley hypothesis can be categorized as being part of first,
artifact-centered approach, to understand anthropomorphism.\\

@@

@@

%{Intro when start with anthropomorphism}
%The essence of anthropomorphism is the perception of human-like characteristics
%in either real or imagined non-human agents. In Human-Robot Interaction (HRI),
%anthropomorphism is a commonly observed phenomenon: people tend to ascribe
%human-like characteristics to robots, and they are willing to treat and react to
%robots as if they were human-social counterparts. The HRI and social robotics
%literature is full of various kinds of studies showing how humans readily
%interact in a social fashion with robots -- despite the fact that robots are not
%humans. In our own studies on HRI analysis in real-world settings, we also
%observed anthropomorphism. However, we found qualitative differences and
%dynamics over time in how participants anthropomorphized and related to the
%robots used in our studies. More concretely, we found that anthropomorphism was
%not only based on how much the robot resembles a human (anthropomorphic design)
%but that it was determined by multiple factors, including individual
%characteristics of the human user but also the context of use, for instance. We also
%found that anthropomorphism changed over time with growing experience of the
%user with the system. By trying to understand these variances and dynamics in
%anthropomorphism, we reviewed related work from the HRI domain and social
%robotics. Despite the fact that there is a vast amount of literature on
%anthropomorphism in HRI, we were surprised to find that other studies do not
%describe similar variances or dynamics. Instead, we came across diverse and
%partially conflicting understandings of anthropomorphism. We also had the
%impression that the reported data were sometimes over-interpreted (probably due
%to the confusion of what exactly anthropomorphism is) or obtained in settings
%and assessed in ways that we believe do not give enough consideration to the
%complex nature of anthropomorphism. However, when reviewing literature from
%social psychology, cognitive science, and neuroscience, we found support for the
%variances and dynamics in anthropomorphism that we observed in our studies. As
%HRI integrates many disciplines, we think that it is valid to use theories from
%these related domains for interpreting our results and that it is relevant to
%report on this integration.
%
%In this article, we would like to share some critical but constructive
%considerations on anthropomorphism in robotics and HRI. Thereby, we contribute
%to a better understanding of this complex phenomenon related to robots. In
%Section \ref{sec:anthropomorphism} we would first like to give an overview of
%how anthropomorphism can be defined, and then, in Section \ref{sec:explanations}
%present explanations to the question why people anthropomorphize, with a focus
%on anthropomorphism as a social phenomenon in HRI and the trend to design
%human-like robots. Then, based on these reviews of theories and concepts from
%related domains, in Section \ref{sec:multi-factors} we propose that
%anthropomorphism is determined and affected by multiple factors. We outline how
%anthropomorphism as a multi-factor phenomenon can impact the user experience and
%acceptance of robots in daily life environments. Next, in Section
%\ref{sec:dynamics-model} we suggest that anthropomorphism changes over time, and
%describe these changes in an initial model that we call ``Dynamics of
%Anthropomorphism''. We illustrate our idea with empirical findings from several
%HRI studies that we conducted in natural environments. In Section
%\ref{sec:cognition-neuroscience} we present a cognitive and neuroscience
%perspective on anthropomorphism to consolidate our considerations and the
%observed dynamics. Next, Section \ref{sec:discussion} considers limitations in
%our approach and we discuss existing and alternative ways to measure
%anthropomorphism in HRI. Finally, in Section \ref{sec:conclusion} we conclude
%this article.

@@

\section{A few results from the field}
\label{sec:overview-studies}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-perception.pdf}

@@

@@

\subsection{A Model of the Dynamics of Anthropomorphism}
\label{sec:dynamics-model}

\paragraph{Why is anthropomorphism likely to change over time?}
\hl{maybe we need to move this paragraph to PART 3}

@@

In this model, anthropomorphism is quantified by a \emph{normalized level of
anthropomorphic effects}: because anthropomorphic effects are not quantified on
an absolute scale, we present them as a normalized value, that spans from a
minimum (no anthropomorphic effects) to a maximum (corresponding to the novelty
effect peak on Figure~\ref{fig:dynamics}). The actual maximum value of
anthropomorphic effects depends on each unique combination of human, robot and
several other factors we introduce below, and thus varies. The general
\emph{shape} of the model remains however the same and depicts the evolution of
anthropomorphism over time, \ie the general dynamics of anthropomorphism.

The model takes into account the duration of the interaction, the nature of the
interaction, as well as acquired experience and familiarization mechanisms. We
also formally introduce a so-called \emph{novelty effect} that models the first
phase of human-robot interaction, during which a specific increase of
anthropomorphic interactions is observed. We focus on \emph{long-term
interaction}, \ie direct (non-mediated), repeated interaction with the same
robot, over an extended period of time (typically longer than a week). The nature 
of the interaction (goal-directed, entertainment, etc.) may
however vary.

@@

\begin{tikzpicture}[scale=1.2]

% background shading
\path[fill=gray!20] (0,0) rectangle (0.6,5.5);
\path[fill=gray!50] (0.6,0) rectangle (5.2,5.5);
\path[fill=gray!20] (5.2,0) rectangle (9.8,5.5);
\draw(0,5.5) node[anchor=south west] {\scriptsize \sc Initialization};
\draw(4,5.5) node[anchor=south] {\scriptsize \sc Familiarization};
\draw(7.5,5.5) node[anchor=south] {\scriptsize \sc Stabilization};
% horizontal axis
\draw[->] (0,0) -- (10,0) node[anchor=north] {$t$};
\draw(5,-0.1) node[anchor=north] {\scriptsize Duration of interaction};

@@

\draw (-0.05, 3) -- (0.05, 3) node[anchor=east] {\scriptsize ICA$_{\mathcal B}$};
\draw (-0.05, 4.1) -- (0.05, 4.1) node[anchor=east] {\scriptsize ICA$_{\mathcal A}$};
\draw (-0.05, 0.6) -- (0.05, 0.6) node[anchor=east] {\scriptsize ICA$_{\mathcal C}$};

% vertical axis - end
\draw[->] (9.8,0) -- (9.8,2) node[anchor=east] {};
\draw (9.8, 0.8) node[anchor=west] {\scriptsize SLA$_\mathcal{E}$};
\draw (9.8, 1.4) node[anchor=west] {\scriptsize SLA$_\mathcal{D}$};
\draw (9.8, .2) node[anchor=west] {\scriptsize SLA$_\mathcal{F}$};

\draw[<-] (0.65,5) -- (0.8,5.2) node[anchor=east] {};
\draw (0.9,5.3) node[anchor=west] {\tiny \it novelty effect};

\draw[dotted] (0, 5) -- (6.2,5);
\draw[dotted] (0, 3) -- (6.2,3);
\draw[<->] (6.1,3) -- (6.1,5) node[anchor=east] {};
\draw (6.2,4) node[rotate=90, anchor=north] {\tiny Expectation mismatch};

\draw[<-] (1.7,4.4) -- (3.2,4.1) node[anchor=east] {};
\draw[<-] (2.85,3.3) -- (3.2,4.1) node[anchor=east] {};
\draw (3.2,4.1) node[anchor=west] {\tiny \it disruptive behaviors};
%%%%%
%% CURVES
%%%%
\begin{scope}[yscale=-1,shift={(-0.125,-0.4)}]

% output of inkscape2tikz
\path[draw=black]
    (0.1250,-2.5620) .. controls (0.1451,-3.7193) and (0.4645,-4.5602) ..
    (0.7044,-4.5633) .. controls (0.8413,-4.5633) and (0.9599,-4.5104) ..
    (1.0794,-4.3971) .. controls (1.1989,-4.2841) and (1.3191,-4.1185) ..
    (1.4588,-3.9181) .. controls (1.5287,-3.8183) and (1.5610,-4.1413) ..
    (1.6381,-4.1420) .. controls (1.7378,-4.1420) and (1.6515,-3.6434) ..
    (1.9554,-3.2280) .. controls (2.1466,-2.9667) and (2.3889,-2.7023) ..
    (2.6819,-2.4292) .. controls (2.7551,-2.3607) and (2.7727,-2.8119) ..
    (2.8771,-2.8180) .. controls (2.9742,-2.8180) and (2.9594,-2.2108) ..
    (3.1665,-2.0209) .. controls (3.3340,-1.8673) and (3.5379,-1.7527) ..
    (3.7508,-1.6236) .. controls (5.8366,-0.4852) and (8.0977,-0.2106) ..
    (10.1250,-0.1860);
\path[draw=black, dashed]
    (0.1250,-3.6924) .. controls (0.1103,-4.0298) and (0.4645,-4.5602) .. 
    (0.7044,-4.5633)
    (3.7508,-1.6236) .. controls (4.9579,-0.9555) and (8.1358,-0.8261) .. 
    (10.1250,-0.7946);
\path[draw=black,dashed]
    (0.1250,-0.1953) .. controls (0.1804,-3.3871) and (0.4645,-4.5602) .. 
    (0.7044,-4.5633) .. controls (0.8413,-4.5633) and (0.9599,-4.5104) .. 
    (1.0794,-4.3971)
    (3.7508,-1.6236) .. controls (4.7654,-0.8505) and (6.5955,0.3538) .. 
    (10.1250,0.4062);

@@

\caption{The dynamics of anthropomorphism. We distinguish three main phases:
    \emph{initialization}, \emph{familiarization} and \emph{stabilization},
    preceded by a \emph{pre-interaction} phase. In the pre-interaction phase,
    users build an \emph{initial capital of anthropomorphism} (ICA). Once the
    interaction starts, the level of anthropomorphism increases due to the
    \emph{novelty effect}~\cite{kanda_interactive_2004}, and then decreases to
    reach a \emph{stabilized level of anthropomorphism} (SLA).  During the
    interaction, unpredicted behaviors of the robot (\emph{disruptive behaviors})
    may lead to local increase of the level of anthropomorphism.}

@@

First, the \emph{initialization} phase. During this short phase (from a couple
of seconds to a couple of hours), we observe an increased level of
anthropomorphism, from an \emph{initial capital of anthropomorphism}
(detailed in the next section) to a peak of anthropomorphic manifestations
that corresponds to the maximum of the \emph{novelty effect}.
Section~\ref{sec:initialization} details this first phase.

@@

The last phase is the \emph{stabilization} phase. The level of anthropomorphic
effects tends to stabilize over a longer time, to reach a \emph{stabilized
level of anthropomorphism} (SLA). The SLA may be null (no anthropomorphic
effects observed anymore), but it may also remain at a higher level.  This
third phase, as well as the \emph{stabilized level of anthropomorphism}, are
discussed in section~\ref{sec:stabilization} below.

@@

The \emph{initial capital of anthropomorphism} describes the initial potential
for the robot to be anthropomorphized by the human user in a given situation.
This potential depends on several factors. It has been shown, for instance, that
some \textit{people} tend to anthropomorphize more than others, that some
\textit{situations} induce anthropomorphism more than others, that
\textit{children} tend to anthropomorphize more than adults, and that some
\textit{cultures} are notorious for their anthropomorphic religions and
worldviews~\cite{epley_when_2008}. Also the shape and design of the robot play a
role, and the context in which the interaction takes place. Our model of
anthropomorphism takes these determinants into account and initializes the level
of anthropomorphic interactions between a human and a robot to a value that we
call \emph{initial capital of anthropomorphism} (ICA). The ICA describes the
first (real or imagined) contact to a robot. In this stage of pre-interaction,
people form initial expectations toward the robot and imagine how they will use
it / interact with it.

We build the ICA on three main factors that \emph{a priori} determine the
potential that a robot will be anthropomorphized:

@@

The ICA comprises some hypothetical values for each of these three factors, such
that different ICA values can be obtained, given the robot, human user, and
environmental context\fxfatal{??}.

For \textbf{personality}, we suggest to apply Epley \textit{et al.}'s
\cite{epley_seeing_2007} psychological \textit{"Three Factor Theory"} of
anthropomorphism. For instance, children generally tend to anthropomorphize
objects more than adults. Also, a person who lacks social connection is said to
be more likely to anthropomorphize. Both aspects would increase the ICA. This
means that, one and the same robot used by a different user, can lead to a
different ICA.

For \textbf{design}, we understand that a robot that follows an anthropomorphic
design (\eg NAO) leads to a higher ICA than a rather functional robot (\eg
Roomba). Also, a robot that is able to display facial expressions would increase
the ICA. However, \hl{as discussed before} a classification of the "amount" of
anthropomorphic design in a robot, as for instance suggested by Fong \textit{et
al.} \cite{fong_survey_2003}, can be difficult. Nevertheless, the important
aspect is how the robot appears to the user, \eg human-like or machine-like.
Studies have shown that an anthropomorphic robot is perceived differently to a
non-anthropomorphic robot\fxfatal{...this sentence sounds strange like that.
Need references or better integration in the paragraph}.

By taking the \textbf{purpose} of a robot into account, we suggest that the real
or imagined context in which a robot is used and the interaction that this usage
brings along, impacts how far the robot will be attributed human-like
characteristics. We draw on findings such as presented in Joosse
\etal~\cite{joosse_what_2013}. The authors showed for instance that when the
same robot (NAO) is used in a different task context (cleaning task \emph{vs.}
tour guide), users ascribe different ``personalities'' to the robot. In general,
a robot which is imagined to be used in a social, entertaining or playful
context leads to a higher ICA than a robot which is used for a routine or
focused task (security, rescue, etc.). This idea also receives support from
Goetz \& Kiesler's work that revealed that people prefer a serious robot for
serious tasks and a less serious robot for more playful
tasks~\cite{goetz_cooperation_2002, goetz_matching_2003}. Also, we suggest that
the environmental context in which people experience and interact with the robot
impacts the ICA. For instance, several friends interacting simultaneously with
the robot might lead to increased ICAs, due to increased human-human social
interactions (the robot might be perceived to be part of the social interaction,
and in turn attributed human-like qualities)~\cite{baxter2013do}.

@@

Our model represents \emph{disruptive behaviors} as local increases of
anthropomorphic effects in the familiarization phase (and at a lesser extend,
in the stabilization phase) (Figure~\ref{fig:dynamics}): because such behaviors
are unexpected, a human observer may interpret them as the result of a richer
deliberative process, which in turn leads to the supposition of complex
cognitive skills

This is however to be modulated depending on the real and percieved reasons of the
unexpected behavior.  Section~\ref{sec:disruptive} discusses more in depth the
mental correlates of disruptive behaviors. It proposes a typology that relates
them to the anthropomorphism they induce. It may increase if the user perceives
(rightfully or not) a form of intentionality, as well as decrease if the
unexpected behavior is perceived as a failure.

@@

The last phase is the \emph{stabilization} phase. The level of anthropomorphic
effects tends to stabilize over a longer time, to reach a \emph{stabilized
level of anthropomorphism} (SLA).

If, after the familiarization phase, no anthropomorphic effects are observed
anymore, the \emph{Stabilized Level of Anthropomorphism} is zero. This can
be interpreted as the user interacting with the robot in a routine way, without
projecting anymore human-like traits on the robot.

\paragraph{Stabilized Level of Anthropomorphism (SLA)}

The \emph{Stabilized Level of Anthropomorphism} (SLA on
Figure~\ref{fig:dynamics}) describes the long-term lasting level of
anthropomorphism.  Like the ICA, the SLA has a unique value per triplet
(\emph{human}, \emph{robot}, \emph{context of interaction}).

We proposed that the ICA is built on three factors: user's \emph{personality},
robot's \emph{design} and interaction \emph{purpose} (or \emph{interaction
context}). The user's personality and the context of use do also influence the
SLA. In particular, it appears that the user's level of acquaintance with
technologies plays an important role in long-term tendency to
anthropomophize~\cite{fink_living_2013} (people more familiar with technology
understand, and hence predict, better the behavior of the robot, which in turn
leads them more frequently to ultimately consider the robot as a simple tool).

The robot's design, on the other hand, plays a more subtle role, and strong
initial anthropomorphic design does not mandate high SLA: lasting
anthropomorphic effects have been observed on non-anthropomorphic robots (like
the iRobot Roomba~\cite{fink_living_2013} or the military iRobot
PackBot\footnote{Rodney Brooks has reported in keynotes that occasionally
soldiers would give a name to \emph{their} PackBot and require it to be repaired
instead of being replaced by another one in case of incident.}), and on the
contrary, anthropomorphic designs can lead to higher expectation deceptions,
resulting in the robot not being used anymore.

Note that the \emph{Initial Capital of Anthropomorphism} and the
\emph{Stabilized Level of Anthropomorphism} are generally not correlated: one
individual may have high potential of anthropomorphizing (high ICA) at first
sight of a good-looking humanoid robot, and get disappointed by the actual
abilities of the robots, down to routine, non-anthropomorphic, interactions (low
SLA), while another user with the same high ICA may, for instance, creates
lasting affective bonds with the same robot, and keeps anthropomorphizing it
(higher SLA).

@@

\subsection{Measuring anthropomorphism}

>>>>

\usepackage{graphicx}
\graphicspath{{figs/}}
\usepackage{tikz}
\usetikzlibrary{shapes}
\usetikzlibrary{svg.path}
%
\usepackage{amsmath} 
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
\usepackage{color, soul}
\usepackage{url}
\usepackage{multirow}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{textcomp}

@@

\usepackage{xifthen}% provides \isempty test

% use either \Ant or \Ant[R,H,C,...]
\newcommand{\Ant}[1][]{%
      \ifthenelse{\isempty{#1}}%
        {$\mathcal{A}$}
        {$\mathcal{A}(#1)$}
}

% for use inside a math environment
\newcommand{\ant}[1][]{%
      \ifthenelse{\isempty{#1}}%
        {\mathcal{A}}
        {\mathcal{A}(#1)}
}
\newcommand{\ICA}{{$\mathcal{A}_0$~}}
\newcommand{\SLA}{{$\mathcal{A}_\infty$~}}
\newcommand{\sla}{{\mathcal{A}_\infty}}
\newcommand{\AntMax}{{$\mathcal{A}_{max}$~}}

@@

A phenomenon that is often studied related to the human-like design of robots is
\textbf{anthropomorphism} (noted \Ant in this paper). Besides a feeling of affinity, and empathy,
anthropomorphism can be understood as being one of the resulting (and  more or
less desired) effects of the anthropomorphic design of artificial agents (up to
the first positive peak). The essence of anthropomorphism is the human user's
perception of human-likeness in the non-human agent. Logically, anthropomorphism
is said to be enhanced with increased human-like design of the agent. There are
various other reasons and explanations for why people tend to anthropomorphize
non-human agents. In general, as anthropomorphism emerges in a (real or
imagined) interaction \citep{persson_anthropomorphism_2000} between at least two
players (the human who anthropomorphizes and the non-human agent which is
anthropomorphized) there are also two main perspectives on the phenomenon: an
artifact-centered (design) and a human-centered (cognitive, psychological) one.
The Uncanny Valley hypothesis can be categorized as being part of first,
artifact-centered approach, to understand anthropomorphism.\\

@@

\subsection{A few results from the field}

To support the reflection, it may be useful to consider a few experimental
results. Figure~\ref{roomba-perception} gives a first perspective on how naive
users perceive robots over a long period of time. Here, \emph{Roomba} robots
were given to nine household for a 6 months period. The aim was to explore how
people use and perceive the robot over time, and how they generally live with
the Roomba in their home. At various time points, users were asked to rated
their perception of different features of the robot on a 7-Likert scale. As
Figure~\ref{roomba-perception} shows, some features remained stable over time
(\emph{ease of use} and \emph{fun}), and other decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one do not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on his/her life. Even if one can argue that the Roomba is not designed
to foster interaction, the question is raised: how to sustain engagement between
a robot and a human over a long period of time? And, as a pre-requisite, do we
actually understand the psychological phenomenon, and can we effectively account
for the dynamics of these interactions?

Those questions are certainly not new, but paradoxically, while anthropomorphism
in robotics is a commonly discussed and studied trait of human-robot
interaction, it appears it lacks a formal ground and unified understanding to
take into account the wide range of phenomena that it encompasses.

\subsection{Towards a formal understanding of anthropomorphism}

So far, the HRI community has not much investigated how anthropomorphism in
human-robot interactions evolves over time (during the process of
\emph{adopting} a robot, for instance). Anthropomorphism is traditionally
perceived in robotics as a static feature that once observed during a
short-term interaction reflects a sustaining social effect.  Based on an
literature review previously published~\cite{fink_anthropomorphism_2012}, a
long-term field study in a natural environment~\cite{fink_living_2013}, as well
as two on-going child-robot experiments, we believe that anthropomorphic
effects do evolve over time in non-monotonic ways, and play a central role in
sustaining engagement in human-robot interaction.

We propose to formalize these dynamic effects into a model that we call the
\emph{dynamics of anthropomorphism}. We hope that it may support further
discussions and reflections on how anthropomorphism impacts human-robot
interaction on the long run, and also foster research on the affective bonds
induced by anthropomorphic projections on robots.

As such, we propose to represent how the level of anthropomorphic effects (\ie
observable manifestations of anthropomorphism) evolves over a long-term
human-robot interaction (Figure~\ref{fig:dynamics}). By long-term interaction,
we mean direct (non-mediated), repeated interaction with the same robot, over
an extended period of time (typically longer than a week).

In this article, we would like to share some critical but constructive
considerations on anthropomorphism in robotics and HRI. Thereby, we contribute
to a better understanding of this complex phenomenon related to robots.

@@

@@

\section{Studying anthropomorphism}
\label{sec:overview-studies}

@@

\begin{figure}
    \centering

@@

The notation \Ant[R,H,C,t] reflects these multiple factors: we suggest here that
anthropomorphism is a function of a vector of robot's traits $R = \{r_0,...,r_n\}$, a vector human
determinants $H = \{h_0,...,h_n\}$, a set of interaction situations that build a
an interaction context $C = \{s_0,...,s_n\}$, and the duration of the
interaction itself, $t$.

@@

@@

While the model primarily considers \emph{long-term interaction} -- direct
(non-mediated), repeated interaction with the same robot, over an extended
period of time (typically longer than a week) --, we also formally introduce a
so-called \emph{novelty effect}~\cite{kanda_interactive_2004} that models the
first phase of human-robot interaction, during which a specific increase of
anthropomorphic interactions is observed.

In this model, anthropomorphism is quantified by a \emph{normalized level of
anthropomorphic effects}: because anthropomorphic effects are not quantified on
an absolute scale, we present them as a normalized value, that spans from a
minimum (no anthropomorphic effects: $\ant=0$) to a maximum (noted \AntMax, with
$t_{max} = \operatorname{arg\,max}_t \, \ant[t]$). We also discuss at
section~\ref{measuring} possible methodologies for the quantitative and
qualitative assessment of anthropomorphic effects.

As pictured on Figure~\ref{fig:dynamics}, we suggest that $t_{max}$ actually
corresponds to the novelty effect peak. The actual maximum value of
anthropomorphic effects depends on each unique combination of human, robot and
several other factors we introduce below, and thus varies. The general
\emph{shape} of the model remains however the same and depicts the evolution of
anthropomorphism over time, \ie the general dynamics of anthropomorphism.

@@

\begin{tikzpicture}[scale=1.4]

% background shading
\path[fill=blue!20] (0,0) rectangle (0.6,5.5);
\path[fill=green!20] (0.6,0) rectangle (5.2,5.5);
\path[fill=orange!20] (5.2,0) rectangle (9.8,5.5);
\draw(0,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Initialization};
\draw(2.5,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Familiarization};
\draw(6.5,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Stabilization};
% horizontal axis
\draw[->] (0,0) -- (10,0) node[anchor=north] {$t$};
\draw(5,-0.1) node[anchor=north] {\scriptsize Duration of interaction};

@@

\draw (-0.05, 3) -- (0.05, 3) node[anchor=east] {\ICA};
\draw (-0.05, 5) -- (0.05, 5) node[anchor=east] {\AntMax};

% vertical axis - end
\draw[->] (9.8,0) -- (9.8,2) node[anchor=east] {};
\draw (9.8, 0.8) node[anchor=west] {\SLA};

\draw[<-] (0.65,5) -- (0.8,5.2) node[anchor=east] {};
\draw (0.9,5.3) node[anchor=west] {\tiny \it novelty peak};

\draw[dotted] (0, 5) -- (6.2,5);
\draw[dotted] (0, 3) -- (6.2,3);
\draw[<->] (6.1,3) -- (6.1,5) node[anchor=east] {};
\draw (6.2,4) node[rotate=90, anchor=north, align=center] {\tiny Expectation\\\tiny mismatch};

\draw[<-] (1.7,4.4) -- (3.2,4.1) node[anchor=east] {};
\draw[<-] (2.85,2.6) -- (3.2,4.1) node[anchor=east] {};
\draw (3.2,4.1) node[anchor=west, align=center] {\tiny \it disruptive\\ \tiny
\it behaviors};
%%%%%
%% CURVES
%%%%
\begin{scope}[yscale=-1,shift={(-0.125,-0.4)}]

    \draw[ultra thick] svg[scale=1cm] "M 0.125,-2.5619582 c 0.0200733,-1.1573591
    0.33954391,-1.9982333 0.57940683,-2.0013597 0.13691605,0 0.25550329,0.052908
    0.37497897,0.1661788 0.119476,0.1130305 0.2396954,0.2786079
    0.37944,0.4790565 0.069872,0.099803 0.1021612,-0.2231749
    0.1792538,-0.2238964 0.099707,0 0.013454,0.4985362 0.3173383,0.9139831
    0.1911917,0.2612926 0.3444354,0.2578391 0.6711932,0.4894291
    0.2058265,0.1458798 0.1571532,0.9762306 0.2615031,0.9702183 0.097182,0
    0.083295,-0.3336208 0.4330446,-0.1199242 0.4306401,0.2631203
    0.216699,0.1355642 0.4296663,0.2647074 2.0858202,1.13836043
    4.34683,1.41300027 6.3741749,1.43753027";

    \draw[dashed] svg[scale=1cm] "M 0.125,-0.1952939 c 0.0553601,-3.1917862
    0.33954391,-4.3648976 0.57940683,-4.368024 0.13691605,0 0.25550329,0.052908
    0.37497897,0.1661788 m 2.6714393,2.7735737 C 4.7653984,-0.85050947
    6.5955093,0.35382293 10.125,0.4062498";

    \draw[dashed] svg[scale=1cm] "M 0.125,-3.6923825 C 0.1102888,-4.0297903
    0.46454391,-4.5601915 0.70440683,-4.5633179 M 3.7508251,-1.6235654 C
    4.9579368,-0.95548345 8.1358175,-0.82609971 10.125,-0.79459549";

@@

\caption{The dynamics of anthropomorphism. We distinguish three main phases:
    \emph{initialization}, \emph{familiarization} and \emph{stabilization},
    preceded by a \emph{pre-interaction} phase. In the pre-interaction phase,
    users build an \emph{initial capital of anthropomorphism} (ICA, noted \ICA).
    Once the interaction starts, the level of anthropomorphism increases due to
    the \emph{novelty effect}, and then decreases to reach a \emph{stabilized
    level of anthropomorphism} (SLA, noted \SLA). \ICA and \SLA may vary
    depending on the user, the robot and the context of interaction.  During the
    interaction, unpredicted behaviors of the robot (\emph{disruptive
    behaviors}) may lead to local increases or decreases of the level of
    anthropomorphism.}

@@

First, the \emph{initialization} phase. During this short phase (from a couple
of seconds to a couple of hours), we observe an increased level of
anthropomorphism, from an \emph{initial capital of anthropomorphism} \ICA
(detailed in the next section) to a peak of anthropomorphic manifestations
\AntMax that corresponds to the maximum of the \emph{novelty effect}.
Section~\ref{sec:initialization} details this first phase.

@@

The last phase is the \emph{stabilization} phase. The level of anthropomorphic
effects tends to stabilize over a longer time, to reach a \emph{stabilized
level of anthropomorphism} \SLA. \SLA may be null (no anthropomorphic
effects observed anymore), but it may also remain at a higher level.  This
third phase, as well as the \emph{stabilized level of anthropomorphism}, are
discussed in section~\ref{sec:stabilization} below.

@@

The \emph{initial capital of anthropomorphism} describes the initial potential
for the robot to be anthropomorphized by the human user in a given situation.
This potential depends on several factors. It has been shown, for instance, that
some \textit{people} tend to anthropomorphize more than others, that some
\textit{situations} induce anthropomorphism more than others, that
\textit{children} tend to anthropomorphize more than adults, and that some
\textit{cultures} are notorious for their anthropomorphic religions and
worldviews~\cite{epley_when_2008}. Also the shape and design of the robot play a
role, and the context in which the interaction takes place. Our model of
anthropomorphism takes these determinants into account and initializes the level
of anthropomorphic interactions between a human and a robot to a value that we
call \emph{initial capital of anthropomorphism}, noted \ICA. \ICA describes the
first (real or imagined) contact to a robot. In this stage of pre-interaction,
people form initial expectations toward the robot and imagine how they will use
it / interact with it.

We build \ICA on three main factors that \emph{a priori} determine the
potential that a robot will be anthropomorphized:

@@

\ICA comprises some hypothetical values for each of these three factors, such
that different \ICA values can be obtained, given the robot, human user, and
environmental context\fxfatal{??}.

For \textbf{personality}, we suggest to apply Epley \textit{et al.}'s
\cite{epley_seeing_2007} psychological \textit{"Three Factor Theory"} of
anthropomorphism. For instance, children generally tend to anthropomorphize
objects more than adults. Also, a person who lacks social connection is said to
be more likely to anthropomorphize. Both aspects would increase \ICA. This
means that, one and the same robot used by a different user, can lead to a
different \ICA.

For \textbf{design}, we understand that a robot that follows an anthropomorphic
design (\eg NAO) leads to a higher \ICA than a rather functional robot (\eg
Roomba). Also, a robot that is able to display facial expressions would increase
the \ICA. However, \hl{as discussed before} a classification of the "amount" of
anthropomorphic design in a robot, as for instance suggested by Fong \textit{et
al.} \cite{fong_survey_2003}, can be difficult. Nevertheless, the important
aspect is how the robot appears to the user, \eg human-like or machine-like.
Studies have shown that an anthropomorphic robot is perceived differently to a
non-anthropomorphic robot\fxfatal{...this sentence sounds strange like that.
Need references or better integration in the paragraph}.

By taking the \textbf{purpose} of a robot into account, we suggest that the real
or imagined context in which a robot is used and the interaction that this usage
brings along, impacts how far the robot will be attributed human-like
characteristics. We draw on findings such as presented in Joosse
\etal~\cite{joosse_what_2013}. The authors showed for instance that when the
same robot (NAO) is used in a different task context (cleaning task \emph{vs.}
tour guide), users ascribe different ``personalities'' to the robot. In general,
a robot which is imagined to be used in a social, entertaining or playful
context leads to a higher \ICA than a robot which is used for a routine or
focused task (security, rescue, etc.). This idea also receives support from
Goetz \& Kiesler's work that revealed that people prefer a serious robot for
serious tasks and a less serious robot for more playful
tasks~\cite{goetz_cooperation_2002, goetz_matching_2003}. Also, we suggest that
the environmental context in which people experience and interact with the robot
impacts the \ICA. For instance, several friends interacting simultaneously with
the robot might lead to increased \ICA for each of the users, due to increased
human-human social interactions (the robot might be perceived to be part of the
social interaction, and in turn attributed human-like
qualities)~\cite{baxter2013do}.

@@

As illustrated on Figure~\ref{fig:dynamics}, our model represents
\emph{disruptive behaviors} as either local increases or local decreases of
anthropomorphic effects in the familiarization phase (and at a lesser extend, in
the stabilization phase): because such behaviors are unexpected, a human
observer may interpret them either as the result of a richer deliberative
process, which in turn leads to the supposition of complex cognitive skills, or
conversely as a failure, reminding the human that the robot is ``just a
machine''. Section~\ref{sec:disruptive} discusses more in depth the impact of
the user's perception of disruptive behaviours, and their cognitive correlates.

@@

The last phase is the \emph{stabilization} phase. The level of anthropomorphic
effects tends to stabilize over a longer time, to reach a \emph{stabilized
level of anthropomorphism}, noted \SLA.

If, after the familiarization phase, no anthropomorphic effects are observed
anymore, $\sla = 0$. This can be interpreted as the user interacting with the
robot in a routine way, without projecting anymore human-like traits on the
robot.

\paragraph{Stabilized Level of Anthropomorphism}

The \emph{Stabilized Level of Anthropomorphism} describes the long-term lasting level of
anthropomorphism.

We proposed that \ICA is built on three factors: user's \emph{personality},
robot's \emph{design} and interaction \emph{purpose} (or \emph{interaction
context}). The user's personality and the context of use do also influence \SLA.
In particular, it appears that the user's level of acquaintance with
technologies plays an important role in long-term tendency to
anthropomophize~\cite{fink_living_2013} (people more familiar with technology
understand, and hence predict, better the behavior of the robot, which in turn
leads them more frequently to ultimately consider the robot as a simple tool).

The robot's design, on the other hand, plays a more subtle role, and strong
initial anthropomorphic design does not mandate high \SLA: lasting
anthropomorphic effects have been observed on non-anthropomorphic robots (like
the iRobot Roomba~\cite{fink_living_2013} or the military iRobot
PackBot\footnote{Rodney Brooks has reported in keynotes that occasionally
soldiers would give a name to \emph{their} PackBot and require it to be repaired
instead of being replaced by another one in case of incident.}), and on the
contrary, anthropomorphic designs can lead to higher expectation deceptions,
resulting in the robot not being used anymore.

Note that the \emph{Initial Capital of Anthropomorphism} and the
\emph{Stabilized Level of Anthropomorphism} are generally not correlated: one
individual may have high potential of anthropomorphizing (high \ICA) at first
sight of a good-looking humanoid robot, and get disappointed by the actual
abilities of the robots, down to routine, non-anthropomorphic, interactions (low
\SLA), while another user with the same high \ICA may, for instance, creates
lasting affective bonds with the same robot, and keeps anthropomorphizing it
(higher \SLA).

@@

\subsection{Measuring anthropomorphism}
\label{measuring}
